---
title: "Sensory priming in aphantasia - Data analysis"
author: "Maël Delem"
format: html
editor_options: 
  chunk_output_type: console
---

Just down below is the setup code, including two essential steps: 1) installing the required packages for this data analysis, and 2) importing data into the dataframes used throughout. I left this step there for reference in a view of transparency for the interested reader.

```{r setup}

# --- Packages -----------------------------------------------------------------

# The package `librairian` will ease the package management with the "shelf" 
# function, which automatically: 
# 1) checks if a package is installed 
# 2) installs it if need be
# 3) loads the package like the "library()" function would.
if (!require(librarian)) install.packages(librarian)
library(librarian)

# now putting packages on our library's shelves:
shelf(
  # Excel files management
  readxl,
  # --- data analysis packages:
  broom,         # tidy modelling
  easystats,      # framework for data analysis functions
  brms,           # bayesian modeling
  corrplot,
  # MASS,           # Box-Cox transformation
  # cluster,        # self-explanatory (i.e. cluster analyses)
  # mclust,         # mixture clustering
  # factoextra,     # multivariate analysis visualization
  # FactoMineR,     # multivariate analysis
  # rstanarm,       # bayesian models
  # BayesFactor,    # self-explanatory (i.e. Bayes Factors)
  # --- data visualization/exploration packages:
  ggpubr,         # layout and statistics display
  # ggradar,        # radar charts
  ggsci,          # scientific palettes
  wesanderson,    # artistic palettes
  viridis,        # colour-blind friendly palettes
  tidygraph,
  ggraph,
  # --- tables ---
  # qwraps2,        # mean_sd function
  # the basics
  tidyverse
)

# --- Global cosmetic theme ----------------------------------------------------
theme_set(
  theme_bw(
    base_size = 14,
    # base_family = "serif",
  )
)

# creating colour-blind friendly coulour palettes
# duo_pal <- c("#003d73", "#889be0")
# trio_pal <- c("#f2bb7b", "#e57774", "#6c0009")
# full_pal <- c("#f2bb7b", "#e57774", "#6c0009", "#003d73", "#889be0")

# table formatting
# options(qwraps2_markup = "")
# options(
#   knitr.kable.NA = '',
#   knitr.table.format = "markdown")

# --- Fixing a seed for reproducibility ----------------------------------------
set.seed(14051998)

# --- Importing data -----------------------------------------------------------

# Association task
df_asso <- read_excel("data/aphantasia_priming_tidy_data.xlsx", 
                     sheet = "data_asso")

# Implicit task
df_implicit <- read_excel("data/aphantasia_priming_tidy_data.xlsx", 
                     sheet = "data_implicit")

# Explicit task
df_explicit <- read_excel("data/aphantasia_priming_tidy_data.xlsx", 
                     sheet = "data_explicit")

# Rotation task
df_rotation <- read_excel("data/aphantasia_priming_tidy_data.xlsx", 
                     sheet = "data_rotation")

# Questionnaires
df_questionnaires <- read_excel("data/aphantasia_priming_tidy_data.xlsx", 
                     sheet = "data_questionnaires")
```

## Exploratory Data Analysis

The four questionnaire variables (VVIQ, OSIQ-O and S, and SUIS) are meant to be used as predictors (independent variables) for the behavioural outcomes (dependent variables, RTs and potentially accuracy) of the study. Three of these - VVIQ, OSIQ-O and SUIS - evaluate visual imagery: as such, in a view of parsimony, we are going to evaluate the correlation between these continuous predictors and judge if they should be merged/omitted. The results of the Bayesian Pearson correlations between the four variables are displayed in @fig-cor_matrix and @tbl-cor_stats.

```{r}
#| label: fig-cor_matrix
#| fig-cap: "Correlation matrix of the questionnaire variables. The stars indicate the amount of evidence in favour of a correlation, as assessed by the $BF_{10}$: No star = No evidence, * = Weak evidence, ** = Moderate evidence, *** = Extreme evidence."

# Bayesian Pearson correlation between questionnaire scores
cor_questionnaires <-
  df_questionnaires %>%
  correlation(
    select = c("vviq80", "suis60","osiq_o75", "osiq_s75"),
    # rename = c("VVIQ", "SUIS", "OSIQ-O", "OSIQ-S"),
    bayesian = TRUE,
    partial = TRUE,
    partial_bayesian = TRUE,
    bayesian_test = "bf"
    )

cor_questionnaires %>%
  summary(digits = 2) %>% 
  plot(
    text = list(size = 5),
    labs = list(title = "")
    ) +
  scale_fill_viridis(
    option = "D",
    name = expression(rho),
    alpha = .6,
    direction = 1,
    limits = c(-1,1)
    ) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 14)
    )
```


```{r}
#| eval: true
#| label: tbl-cor_stats
#| tbl-cap: "Detailed correlation matrix of the questionnaire variables."

cor_questionnaires %>%
  format %>% 
  rename(
    "Variable 1" = Parameter1,
    "Variable 2" = Parameter2,
    "$\\rho$" = rho,
    "$BF_{10}$" = BF
  ) %>% 
  knitr::kable()
```






























